Manuscript Preparation

### 2018-06-20 Houston/src/manuscript_preparation.Rmd
Add ensemble learning

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
options(scipen = 1, digits = 2)
require(dplyr)
require(ggplot2)
require(RcppRoll)
require(markovchain) # FIX: markovchain seems not compatible with R3.5.0
require(reshape)
require(survival)
require(caret)
require(e1071)
require(randomForest)
# knitr::opts_knit$set(root.dir = "~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
# source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/prepare_violation.R")
# source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/roll_over.R")
#knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN

@@

@@

Next, let's try out spectral clustering.  (TODO)

@@

```{r randomForest, include=FALSE}
# reformat labeled.data.lost.train$severe to factors so that `randomForest`
# can do a classification
labeled.data.lost.train$severe <- labeled.data.lost.train$severe %>% factor
labeled.data.lost.result$severe <- labeled.data.lost.result$severe %>% factor

@@

Now, let's try an ensemble learning method, which is a combination of random forest and SVM.

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
options(scipen = 1, digits = 2)
require(dplyr)
require(ggplot2)
require(RcppRoll)
require(markovchain) # FIX: markovchain seems not compatible with R3.5.0
require(reshape)
require(survival)
require(caret)
require(e1071)
require(randomForest)
require(kernlab)
require(SuperLearner)
# knitr::opts_knit$set(root.dir = "~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
# source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/prepare_violation.R")
# source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/roll_over.R")
#knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN

@@

#save(clustering, file="kmeans_clustering.RData")

load("./Houston/output/kmeans_clustering.RData")

@@

Next, let's try out spectral clustering, which is based on the connectivity between data points.

```{r sprectral clustering, include=FALSE}
temp <- complete.active.quarters %>% filter(active==TRUE)
set.seed(20)

sc.model <- specc(temp[, 6:ncol(temp)], centers=3)
#save(sc.model, file="sc_model.RData")

load("./Houston/output/sc_model.RData")

#plot(temp[, 6:ncol(temp)], col=sc.model, pch=4) # too many features, very slow
plot(temp[, 6], col=sc.model, pch=4) # plot num.days.lost cluster

The dataset is highly skewed, so K-means does not perform well as expected since it uses means to estimate centers. On the other hand, spectral clustering uses affinity between data points to cluster. However, the performance is poor as well.

@@

```{r randomForest, include=FALSE}
# reformat labeled.data.lost.train$severe to factors so that `randomForest`
# can do a classification
labeled.data.lost.train$severe <- labeled.data.lost.train$severe %>% factor
labeled.data.lost.result$severe <- labeled.data.lost.result$severe %>% factor
set.seed(20)

@@

@@

Now, let's try an ensemble learning method, which is a combination of single learning algorithms. In this case, let's try svm and random forest and see if the ensemble learning could improve the performance.

Read [this post](https://www.datacamp.com/community/tutorials/ensemble-r-machine-learning) first.

```{r ensemble, include=FALSE}
# encode response variable for superlearner
y <- as.numeric(labeled.data.lost.train$severe)-1
ytest <- as.numeric(labeled.data.lost.result$severe)-1

# separate predictors
x <- data.frame(labeled.data.lost.train[, c('last.quarter.lost','last.year.lost',
                                            'last.three.years.lost','last.quarter.restrict',
                                            'last.year.restrict', 'last.three.years.restrict',
                                            'last.quarter.viol','last.year.viol',
                                            'last.three.years.viol', 'last.quarter.death',
                                            'last.year.death', 'last.year.death',
                                            'last.three.years.death')])
xtest <- data.frame(labeled.data.lost.result[, c('last.quarter.lost','last.year.lost',
                                            'last.three.years.lost','last.quarter.restrict',
                                            'last.year.restrict', 'last.three.years.restrict',
                                            'last.quarter.viol','last.year.viol',
                                            'last.three.years.viol', 'last.quarter.death',
                                            'last.year.death', 'last.year.death',
                                            'last.three.years.death')])

# TODO: 1. try default params first
#       2. then try tuned params with ksvm and randomForest
#       3. try some other learners such as glmnet
ensemble.model <- SuperLearner(y,
                               x,
                               family=binomial(),
                               SL.library=list("SL.randomForest", # TODO: use ranger, SL.randomForest will be removed from SuperLearner
                                               "SL.ksvm") 
                               )

SL.randomForest.tune <- function(...){
      SL.randomForest(..., num.trees=200, mtry=2, nodeside=240)
    }

ensemble.model.tune <- SuperLearner(y,
                                    x,
                                    family=binomial(),
                                    SL.library=list("SL.ranger", # TODO: do `install.packages("ranger")` first if using `SL.ranger`
                                                    "SL.ksvm",
                                                    "SL.randomForest.tune") 
                                    )

ensemble.model.more <- SuperLearner(y,
                                    x,
                                    family=binomial(),
                                    SL.library=list("SL.ranger", # TODO: need pacakge `ranger`, please do `install.packages("ranger")`
                                                    "SL.ksvm",
                                                    "SL.randomForest.tune",
                                                    "SL.glmnet") # TDOD: need package `glmnet`
                                    )
